---
title: "class_12a-asking for help in R"
author: "Brandon Brenes"
date: "2024-11-19"
output: html_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

```

**Be specific,**

-   Search with the keywords and package/function name if known, and include .R in the search keywords (e.g. “dplyr::mutate add column in .R” instead of “mutate variable”)

-   “R” is generic - so consider using “R software” or “.R” Start learning and using language common in R communities & publications (e.g. R4DS), like data frame or tibble instead of spreadsheet

-   For now, anything reading, wrangling, ggplot2 related: considering navigating to documentation from [tidyverse.org](https://www.tidyverse.org/), [rdocumentation](https://www.rdocumentation.org/), or [RStudio/Posit community](https://community.rstudio.com/).

**If you’re asking: What package or function should I use to do this thing?**

-   Don’t know what package/function to use for your purpose? Use [CRAN Task Views](https://cran.r-project.org/web/views/) to help you find it (grouped by topic) to search for packages by keyword

**A lot of these** **errors can be pointed out with chat and or pre-installed R warnings/ERRORS.**

**How to find out what arguments are accepted by which functions:**

-   Viewing the R documentation with ?function_name, and look in the
    ‘Arguments’ section (or ‘Aesthetics’ section for geoms in ggplot2)


```{r}

# libraries #

library(styler)
library(reprex)
library(datapasa)
library(tidyverse)

```

```{r}
library(tidyverse)
mpg %>%
  ggplot(aes(x = displ, y = hwy))%>%
  geom_point(aes(color = class))

```

Let's say you are trying to make a plot with the mpg dataset and you hit an error.

```{r}

mpg %>%
  ggplot(aes(x = displ, y = hwy))%>%
  geom_point(aes(color = class))

```

Go to Addins -> Render Reprex -> Paste to the program you want.
Make sure to "append session info"

It formats it for you!

Another example...

Using real data...
lat    long    star_no
33.548    -117.805    10
35.534    -121.083    1
39.503    -123.743    25
32.863    -117.24    22
33.46    -117.671    8
33.548    -117.805    3

this is useless^^

highlight/copy directly from csv and paste as a tribble in the addins section...
```{r}
stars<-tibble::tribble(
      ~lat.long....star_no,
  "33.548  -117.805    10",
   "35.534  -121.083    1",
  "39.503  -123.743    25",
      "32.863  -117.24 22",
   "33.46   -117.671    8",
   "33.548  -117.805    3",
  "33.603  -117.879    15",
  "34.39   -119.511    23",
   "35.156  -120.674    4",
   "35.316  -120.845    7",
  "35.534  -121.083    35",
  "36.623  -121.919    21",
  "38.306  -123.014    14",
  "41.78   -124.234    17"
  )


```


```{r}

library(forcats)
library(tidyverse)

# Separate columns into lat, long, and star_no
star <- star %>%
  separate(
    lat_long_star_no, 
    into = c("lat", "long", "star_no"), 
    sep = "\\s+"
  ) %>% 
  mutate(star_no = as.numeric(star_no)) #ensure star_no is numeric

# Reorder rows by star_no
star <- star %>%
  rearrange(star_no)




```




