---
title: "13a_class"
author: "Brandon Brenes"
date: "2024-12-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# libraries

```{r}
library(tidyverse)
library(here)
```

**for loops** are one way to run an iterative process (do something over and over again). However, with the tidyverse you can avoid needing loops for almost any basic coding needs. Where it does come in handy is in population modeling and Bayesian modeling, for example. Also, it is integral to programming and is happening "under the hood" whether you use them or not.

```{r}
years<-c(2015:2021)

for (i in years) {
  print(paste("The year is", i))
  
}
```


Add the for loop

One line at a time. Let's first add in the column that is going to have all the names in it. Notice that I added an index i in the column name. I also am having the index go from 1:length(years), which is 1:7. I use length() because it allows me to change the number of years in the vector without having to change the for loop. 

```{r}

#this is the broadest way to make loops to allocate space:

#Pre-allocate space for the for loop
# empty matrix that is as long as the years vector

year_data<-tibble(year =  rep(NA, length(years)),  
 # column name for year
                  year_name = rep(NA, length(years))) 
# column name for the year name
year_data

```

```{r}

for (i in 1:length(years)){ # set up the for loop where i is the index
  year_data$year_name[i]<-paste("The year is", years[i]) # loop over i
}
year_data
```
```{r}
testdata<-read_csv(here("Week_13", "data","cond_data","011521_CT316_1pcal.csv"))

glimpse(testdata)

```

# Using for loop

```{r}
# point to the location on the computer of the folder
CondPath<-here("Week_13", "data", "cond_data")
# list all the files in that path with a specific pattern
# In this case we are looking for everything that has a .csv in the filename
# you can use regex to be more specific if you are looking for certain patterns in filenames
files <- dir(path = CondPath,pattern = ".csv")
files


cond_data<- tibble(filename =  rep(NA, length(files)),  # column name for year
                   mean_temp = rep(NA, length(files)), # column name for the mean temperature
                   mean_sal = rep(NA, length(files)), # column name for the mean salinity
                   ) # column name for the year name


raw_data<-read_csv(paste0(CondPath,"/",files[1])) # test by reading in the first file and see if it works

head(raw_data)


for (i in 1:length(files)){ # loop over 1:3 the number of files
}

for (i in 1:length(files)){ # loop over 1:3 the number of files 
raw_data<-read_csv(paste0(CondPath,"/",files[i]))
glimpse(raw_data)
}

for (i in 1:length(files)){ # loop over 1:3 the number of files 
raw_data<-read_csv(paste0(CondPath,"/",files[i]))
#glimpse(raw_data)
cond_data$filename[i]<-files[i]
} 

cond_data


for (i in 1:length(files)){ # loop over 1:3 the number of files 
raw_data<-read_csv(paste0(CondPath,"/",files[i]))
#glimpse(raw_data)
cond_data$filename[i]<-files[i]
cond_data$mean_temp[i]<-mean(raw_data$Temperature, na.rm =TRUE)
cond_data$mean_sal[i]<-mean(raw_data$Salinity, na.rm =TRUE)
} 

cond_data
```

 # simpler ways to do this using map() function
 
1. Use a canned function that already exists...

```{r}
1:10 # a vector from 1 to 10 (we are going to do this 10 times)

1:10 %>% # a vector from 1 to 10 (we are going to do this 10 times) %>% # the vector to iterate over
  map(rnorm, n = 15) # calculate 15 random numbers based on a normal distribution in a list

```
Another example...

```{r}
1:10 %>% # a vector from 1 to 10 (we are going to do this 10 times) %>% # the vector to iterate over
  map(rnorm, n = 15) %>% 
map_dbl(mean) # calculate the mean. It is now a vector which is type "double"
```


2. Make your own function...

```{r}
1:10 %>% # list 1:10
  map(function(x) rnorm(15, x)) %>% # make your own function
  map_dbl(mean)
```
3. Canned function but changing the argument with the function

```{r}
1:10 %>%
  map(~ rnorm(15, .x)) %>% # changes the arguments inside the function
  map_dbl(mean)
```

```{r}
CondPath<-here("Week_13", "data", "cond_data")

files <- dir(path = CondPath,pattern = ".csv")

files

# or we can do this if we want the full name

files <- dir(path = CondPath,pattern = ".csv", full.names = TRUE)
#save the entire path name

files



data<-files %>%
  set_names()%>% # set's the id of each list to the file name
  map_df(read_csv,.id = "filename") # map everything to a dataframe and put the id in a column called filename

data

```

 