---
title: "Markdown HW for Week 6"
author: "Brandon Brenes"
date: "2024-10-05"
output:
  html_document:
    toc: true
    toc_float: true
  pdf_document:
    toc: true
    css: styles.css  #CSS file for custom styles
    
---  
  
____________
  
   
   
<br>
     
     
     
## **Description :**
#### This markdown file is a homework assignment for OCN 682 to practice my skills in _Rmarkdown_. 

#### We will be creating a clean, easy-to-read **Markdown** document that utilizes the chemical data from Maunalua. The data will be displayed below in three formats: 

* Raw data
* Table
* Heatmap plot
  
<br>
  
## 1. Libraries
#### These are the libraries being used to process this data:

<br>

```{r setup, message=FALSE}
library(gt)
library(here)
library(readr)
library(dplyr)
library(tidyr)
library(shiny)
library(tinytex)
library(ggplot2)
library(gtExtras)
library(tidyverse)
library(kableExtra)
library(wesanderson)
```

<br>

## 2. Raw Data
#### We'll now be looking at the **chemistry data from Maunalua**. The data was read in as a csv file using _read_csv()_. Here is a glimpse of the raw data:

<br>

```{r, echo=FALSE, message=FALSE}

# read in data #

chemical_data <- readr::read_csv("/Users/brenesb/Desktop/Repositories/Brenes/Week_06/data/chemicaldata_maunalua.csv")

glimpse(chemical_data) #display
 
heatmap_data <- chemical_data %>% #rename#convert from wide to long format for the heat map
  select(Zone, Season, Phosphate, Silicate, NN) %>% #select for only these columns
  pivot_longer(cols = c(Phosphate, Silicate, NN), #pivot to ease heatmap creation
               #pivoted to long format to make heat map
               names_to = "Nutrient", #names go to Nutrient column
               values_to = "Concentration") #values are in concentration column
```

<br>

## 3. Summary Table
#### This is a table showing the **mean concentration (umol/l)** with **varying zones** and **seasons**. The data was grouped by zone, season and nutrients, then pivoted wider to make it more readible. I used the New York Times theme to give it a clean and simple look.

<br>

```{r summary, echo=FALSE, message=FALSE,fig.show='hold'}

# summary table #

summary_table <- heatmap_data %>%
  group_by(Zone, Season, Nutrient) %>% #group by zone, season and nutrient
  summarise(Mean_Concentration = mean(Concentration, na.rm = TRUE), .groups = 'drop') %>% #find average conc. of nutrients
  drop_na() %>% #get rid of NAs
  pivot_wider(names_from = Nutrient, values_from = Mean_Concentration) %>% #pivot to improve readbility 
  rename('Nitrate + Nitrite (umol/l)' = NN) %>% #rename to include 'umol/L'
  rename('Silicate (umol/l)' = Silicate) %>%
  rename('PHOSPHATE (umol/l)' = Phosphate) %>%
  gt() %>% #gt function used for NYT theme
  gt_theme_nytimes() %>% #utilize NYT theme
  tab_header(title = "Nutrients Across Space and Time") #title of table

# display #
summary_table

# save table as png #
gtsave(summary_table, here("Week_06","output", "summary_table.png"))

```

<br>

## 4. Heatmap Nutrient Plots
#### The raw data was pivoted longer to make it easier to plot a heatmap using _geom_tile()_ from ggplot.

<br>

```{r heatmap, echo=FALSE, message=FALSE, fig.show='hold'}

# for silicate #

silicate_map <- ggplot(heatmap_data %>%
                         drop_na()%>%
                         filter(Nutrient == "Silicate"), #filter for only silicate
                       aes(x = Zone, y = Season, fill = Concentration)) +#set x and y to zone and season, fill will be conc. of the nutrient in umol/L
  geom_tile() + #heatmap using geomtile
  scale_fill_gradientn(colors = colorRampPalette(c("white", "blue"))(100), na.value = "grey50") +
  labs(title = "Silicate", fill = "umol/L") +#title for plot and legend
  theme_bw() +
  theme(panel.grid.major = element_blank(),  #remove ugly grid lines that show up on plot
        panel.grid.minor = element_blank(),
        plot.title = element_text(size = 16, face = "bold"), #adjust size and bold of all titles
    axis.title.x = element_text(size = 14, face = "bold"),
    axis.title.y = element_text(size = 14, face = "bold"),
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 12),
    
  )

# display #
print(silicate_map)

# save plot #
ggsave(here("Week_06","output","silicate_plot.png"), width = 12, height = 8)
```

<br>

```{r heatmap2, echo=FALSE, message=FALSE, fig.show='hold'}

# for nitrate + nitrite #

nn_map <- ggplot(heatmap_data %>%
                   drop_na()%>%
                   filter(Nutrient == "NN"), #filter for only NN
                 aes(x = Zone, y = Season, fill = Concentration)) +#set x and y to zone and season, fill will be conc. of the nutrient in umol/L
  geom_tile() + #heatmap using geomtile
  scale_fill_gradientn(colors = colorRampPalette(c("white", "red"))(100), na.value = "grey50") +
  labs(title = "Nitrate + Nitrite", fill = "umol/L") + #title for plot and legend
  theme_bw() +
  theme(panel.grid.major = element_blank(),  #remove ugly grid lines that show up on plot
        panel.grid.minor = element_blank(),
        plot.title = element_text(size = 16, face = "bold"), #adjust size and bold of all titles
    axis.title.x = element_text(size = 14, face = "bold"),
    axis.title.y = element_text(size = 14, face = "bold"),
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 12)
  )

# display #
print(nn_map)

# save plot #
ggsave(here("Week_06","output","nn_plot.png"), width = 12, height = 8)
```

<br>

```{r heatmap3, echo=FALSE, message=FALSE, fig.show='hold'}

# for phosphate #

phosphate_map <- ggplot(heatmap_data %>% 
                          drop_na()%>%
                          filter(Nutrient == "Phosphate"), #filter for only phosphate
                         aes(x = Zone, y = Season, fill = Concentration)) + #set x and y to zone and season, fill will be conc. of the nutrient in umol/L
  geom_tile() + #heatmap using geomtile
  scale_fill_gradientn(colors = colorRampPalette(c("white", "green"))(100), na.value = "grey50") +
  labs(title = "Phosphate", fill = "umol/L") + #title for plot and legend
  theme_bw() +
  theme(panel.grid.major = element_blank(),  #remove ugly grid lines that show up on plot
        panel.grid.minor = element_blank(),
        plot.title = element_text(size = 16, face = "bold"), #adjust size and bold of all titles
    axis.title.x = element_text(size = 14, face = "bold"),
    axis.title.y = element_text(size = 14, face = "bold"),
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 12)
  )

# display #
print(phosphate_map)

# save plot #
ggsave(here("Week_06","output","phosphate_plot.png"), width = 12, height = 8)
```

<br>

###### last updated: 2024/10/06
