---

title: "week6a"
author: "Brandon Brenes"
date: "2024-10-01"
output:
  html_document:
    toc: TRUE
    toc_float: TRUE
    theme: darkly
---
# General Notes #

hit space bar twice and then return to make a new line

Add R chucks, chunks of area for Rcoding by hitting the "+C" above and hitting R or by typing...

```{r setup, include=FALSE}
```


There are 'chunk options': getting rid of warnings, add a caption to graphical resutls, etc. Include= false 
means no code will show up

```{r setup, include=FALSE, echo = TRUE, message = FALSE, warning = FALSE}
```

# Intro to Penguins:

Today we are going to work with **penguin data** in _Markdown_

# load libraries:
```{r setup, include=FALSE}

library(palmerpenguins)
library(tidyverse)
library(here)
```

# inspect the data:
```{r}
glimpse(penguins)
```

# make a plot
```{r}
plot1<-ggplot(data = penguins, 
              mapping = aes(x = bill_depth_mm, 
                            y = bill_length_mm,
                            group = species,
                            color = species)) +
  geom_point() + 
  geom_smooth(method = "lm") + 
  labs(x = "Bill Depth (mm)", y = "Bill Length (mm)") + 
  scale_x_continuous(breaks = c(14, 17, 21), #this adds breaks to your x axes
                     labels = c("low", "medium", "high"))+
  theme_classic() +
  theme(axis.title = element_text(size = 22, color = "red"),
        panel.background = element_rect(fill = "linen"))
```

