---
title: "class_11a"
author: "Brandon Brenes"
date: "2024-11-12"
output: html_document
---

## Description

Intro to shiny app

-   What is Shiny and why use it?

-   Basics of user interface

-   Basics of the server

-   Reactive objects

-   Resources for Shiny Themes, dashboard, publishing your app

Shiny is used as a middle man between R and website code interface,etc.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## libraries

```{r}

# libraries #

library(shiny)

```

### NOTES on Shiny R

User interface (ui) always starts with fluidPage(), server() function will be the background of the code

The **`ui`** (user interface) object dictates the appearance of the app. UI functions write HTML. For something to appear in your app, it needs to be in the UI.

The **`server()`** function contains rendering expressions that create the objects to display.

The server function and UI object are passed to the `shinyApp()` function to create a Shiny app object.

**When saving the script in your app, create a folder with the name of your app and then save the script as 'app.R'. You MUST save the script with this name for it to work. Also, all files associted with your app must be in this folder.**\

We are working within app.R in ClassApp...
