---
title: "Tidy Tuesday Volcano"
author: "Brandon Brenes"
date: "2024-11-08"
output:
  html_document:
    toc: true
    toc_float: true
    theme:
      bootswatch: flatly
    highlight: tango
pdf_document:
  toc: true
  highlight: tango
  latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

## Description

We will be making a *geom_map* plot using *geom_polygo*n and *geom_point* displaying the **volcanoes in Indonesia**, the volcanoes **tectonic setting** and **elevation**.

<br>

## Libraries

```{r}

# libraries #

library(here)
library(maps)
library(readr)
library(dplyr)
library(bslib)
library(mapdata)
library(mapproj)
library(tidyverse)

```

<br>

## Data

### Load data

The volcano data came from a *tiny_tuesday* repository named **Volcano Eruptions.** The world map data came from the **maps package**

```{r}

# load data #

volcano <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-05-12/volcano.csv')

world<-map_data("world")

# display #
head(world)

# display #
head(volcano)

```

<br>

### Data manipulation

First we will filter for Indonesia lat and long data and filter volcano data to only have a necessary variables and only Indonesia data.

```{r}

# world data manipulation #

indonesia <- world %>% filter(region == "Indonesia") 
#filter for only Indonesia data


# volcano data manipulation #

indonesia_volcano_subset <- volcano %>%
  select(volcano_number, country, tectonic_settings, latitude, longitude, elevation) %>%
  filter(country == "Indonesia")
#filter for only Indonesia data and necessary columns


# display #

head(indonesia_volcano_subset)

```

<br>

## Plot

Using the *indonesia* with the **geom_polygon()** function and *indonesia_volcano_subset* data with the **geom_point** function overlay-ed, I created a map of volcanoes in Indonesia. I then used the **scale_size_continuous** and **labs()** function to express the relationship between the size of the point and the elevation of the volcano The color was filled using the tectonic setting as shown in the legend.

```{r indonesia-volcano-plot, echo=TRUE, fig.width=20, fig.height=15, dpi=100}

indonesia_volcano_plot <- ggplot() +
    #plot the shape of Indonesia
    geom_polygon(data = indonesia, 
                 aes(x = long, y = lat, group = group),
                 fill = "beige", color = "#BC8F8F") + #beige and brown
 #add points for each volcano, with size based on elevation and color by tectonic_settings
    geom_point(data = indonesia_volcano_subset, 
               aes(x = longitude, y = latitude, size = elevation, color = tectonic_settings),
               alpha = 0.5) + #add transparency for better visibility
    #customize the size scale for elevation
    scale_size_continuous(
        name = "Elevation (m)", 
        breaks = c(500, 1000, 3000), 
        range = c(0.5, 4)  # increase the range for bigger points
    ) +
    #use a mercator projection
    coord_map(projection = "mercator", xlim = c(95, 150), ylim = c(-15, 15)) +
    theme_light()+
    scale_color_brewer(name = "Tectonic Setting", palette = "Dark2")+
    labs(title = "Volcanoes in Indonesia by Elevation and Tectonic Setting")+
    #customize background color and layout
  theme(
        plot.title = element_text(face = "bold", hjust = 0.1, size = 35, margin = margin(b = 20)),  
        #bold title, left aligned
        plot.margin = margin(t = 10, r = 10, b = 10, l = 10),  
        #add margin for spacing
        panel.background = element_rect(fill = "lightblue2"),  
        #set background of plot area
        legend.background = element_blank(),  
        #remove background from the legend area
        legend.key = element_blank(),  
        #make the legend keys (symbols) transparent
        axis.text = element_blank(),  #remove axis labels and grid
        axis.ticks = element_blank(),  
        axis.title = element_blank(),  
        panel.grid = element_blank(),  
        legend.title = element_text(size = 20, face = "bold"), #legend title
        legend.text = element_text(size = 20),#legend text
        legend.spacing = unit(0.2, "cm"),  #reduce space between the legends
        legend.box.spacing = unit(0.2, "cm")  #reduce space between legends
    )

indonesia_volcano_plot

```

## What did I learn?

I learned how to make a ggmap displaying **two variables** using the color and the point size to represent different values. The colors display an interesting pattern showing the longitudinal change in tectonic setting. I am also using a color blind friendly pallet, which is something that I've been meaning to make a habit in my research analyses. I also used a new rmarkdown theme!
