---
title: "Tidy Tuesday Orca"
author: "Brandon Brenes"
date: "2024-11-02"
output:
  prettydoc::html_pretty:
    theme: cayman
    toc: true
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

```

<br>

# Description

We will be looking to see if there is a seasonal variation in the encounter number at Haro Strait. To do this we will make a violin plot with season on the x axis and encounters on the y axis. The challenge for me will be to learn how to make seasonal data using their *date* variable. In addition to this, I will also be making a violin plot for the first time and utilize a new theme for rmarkdown and ggplot.

<br>

# Libraries

```{r}

# library #

library(tidyr)
library(dplyr)
library(readr)
library(ggplot2)
library(cowplot)
library(forcats)
library(prettydoc)
library(lubridate)

```

<br>

# Load Data

```{r}

# load data from tidy tuesday orcas #

raw_orcas <-read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2024/2024-10-15/orcas.csv')

head(raw_orcas)


```

<br>

# Data manipulation

**Clean up the data:**

Select on the necessary columns, filter data that was collected in Haro Strait and create a season column using the dates (lubridate).

```{r}

# clean up raw data #

# select out the columns we need (date, year and encounter) and only data in location Haro Strait, manipulate the data give seasons for dates.

clean_orcas <- raw_orcas %>% 
  filter(location == "Haro Strait") %>% #only data from Haro Strait
  select(year, date, encounter_number) %>% #select out variables needed
  mutate(
    date = ymd(date),
    season = case_when(#make month within these numbered months equal to this "season"
      month(date) %in% 3:5  ~ "Spring",
      month(date) %in% 6:8  ~ "Summer",
      month(date) %in% 9:11 ~ "Fall",
      TRUE                  ~ "Winter"
    )
  )

head(clean_orcas)
```

<br>

**Summary data for plot:**

Find the difference between encounter numbers in each season because the number encounter number restarts every year.

```{r}

summary_orcas<-clean_orcas %>%
 group_by(year, season) %>% #group by year and season
  summarize(#subtract the min from the max for ecounter number for that years seasons.
    min_encounter = min(encounter_number, na.rm = TRUE),
    max_encounter = max(encounter_number, na.rm = TRUE),
    encounter_diff = max_encounter - min_encounter
  ) %>%
  ungroup() %>% #ungroup just for safety reasons
  drop_na() #drop na from dataset

head(summary_orcas)

```

<br>

# Plot

Now we use the encounter diff as the number of encounters in that season and create a violin plot showing the seasonal variation in encounters:

```{r}

# violin plot of number of occurrences by season

orca_plot<- ggplot(summary_orcas, aes(x = fct_relevel(season, "Spring", "Summer", "Fall", "Winter"), y = encounter_diff)) +#use fct_relevel so it puts seasons in order
  geom_violin(aes(fill = season)) + #violin plot
  labs(
    x = "Season",
    y =  "Number of Encounters",
    title = "Orca Encounter Numbers by Season in Haro Strait",
    fill = "Season"
  ) +
  theme_minimal_grid() + #new theme!
  theme(
    legend.position = "none",                # removes the legend
    axis.title = element_text(face = "bold") # bold axis titles
  )


library(cowplot)

orca_plot <- ggplot(summary_orcas, aes(x = fct_relevel(season, "Spring", "Summer", "Fall", "Winter"), y = encounter_diff)) +
  geom_violin(aes(fill = season)) + 
  labs(
    x = "Season",
    y =  "Number of Encounters",
    title = "Orca Encounter Numbers by Season in Haro Strait",
    fill = "Season"
  ) +
  theme_minimal_grid() +  # Start with theme_minimal_grid from cowplot
  theme(
    panel.background = element_rect(fill = "white", color = NA),  # White background
    plot.background = element_rect(fill = "white", color = NA),   # White background around plot
    legend.position = "none",                                     # Removes the legend
    axis.title = element_text(face = "bold")                      # Bold axis titles
  )

orca_plot

```

<br>

# What did I learn?
In this week's _tidy_tuesday_, I learned how to make seasons out of dates - a very useful tool when looking at time-series data. I also learned how to make a violin plot, which perfectly displayed the spread in values for these encounters over time. Lastly, I used different themes for the plot and the rmarkdown html file, both of which had to be installed.
