---
title: "Presidents Age Tidy Tuesday"
author: "Brandon Brenes"
date: "2024-11-25"
output:
  html_document:
    theme: united
    toc: true
    toc_float: true
    highlight: zenburn
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

```

<br>

# Description

We are going to be making a box plot of the various president ages on the y axis and the various countries (US, UK and Japan) on the x axis. To improve readability, superimpose some jitter points on the plot.

<br>

# Libraries and Load Data

```{r}

# library #

library(tidyr)
library(dplyr)
library(here)
library(readr)
library(ggsci)
library(ggplot2)
library(kableExtra)
library(RColorBrewer)

```

```{r}

# load data from tidy tuesday democracy data #

raw_democracy <- read_csv(here("tidy_tuesday","data", "w13_democracy_data.csv"))

# display #
head(raw_democracy)

```

<br>

# Data manipulation

**Clean up the data**

We are going to select out the columns we need (president_brithdyear, president_accession_year and country_name) then group by country_name.

```{r}

# clean up raw data #

# select for president birthyear, acccesion year and country #

clean_democracy <- raw_democracy %>% #filter/select out important variables 
  select(president_birthyear, president_accesion_year, country_name) %>%
  group_by(country_name) %>%  #group by country
  drop_na() #drop nas

clean_democracy %>% 
  head(6) %>% 
  kable() 

```

<br>

**Summary data for plot**

We will now make an age column by finding the difference of the two year variables. This will allow us to have a data set we can make an age vs. country box plot

```{r}
# summary data #

# find the age of the president at the time he/she were elected #

summary_democracy <- clean_democracy %>%
  mutate(president_age = president_accesion_year - president_birthyear) %>% 
   # calculate and add the president age
  filter(country_name %in% c("Indonesia", "United States", "France")) %>%
  group_by(country_name) # group by country for plot

# display #
summary_democracy %>% 
  head(6) %>% 
  kable() 

```

**Mean table**

**Calculate mean to visualize data in table format**

```{r}

# calculate mean to visualize data in table format #

# table mean data peak #
mean_age_by_country <- summary_democracy %>%
  group_by(country_name) %>%
  summarise(mean_age = mean(president_age, na.rm = TRUE))

# display #
mean_age_by_country %>% 
  head(6) %>% 
  kable() 

```

<br>

# Plot

Using ggplot and their functions geom_boxplot and geom jitter, I made a boxplot showing the spread and median of president ages in the countries France, United States and Indonesia. I used a scientific color pallete, which I hadn't heard of before. Very clean and appealing look.

```{r}

# create the box plot and jittered individual points using ggplot geom_jitter and geom_boxplot #

president_age_box_plot<-ggplot(summary_democracy, aes(x = country_name, y = president_age, fill = country_name)) +
  geom_boxplot(outlier.shape = NA) +  # hide outliers from boxplot
  geom_jitter(width = 0.2, alpha = 0.5, size = 2) +  # add jittered individual points with transparency
  scale_fill_npg() +  #npg pallete
  labs(title = "President Age Distribution by Country",
       x = "Country",
       y = "President Age") +
  theme_light() +  # light theme for the plot
  theme(
    plot.title = element_text(face = "bold", size = 20),  # bold plot title
    axis.title = element_text(face = "bold"),  # bold 
    axis.text = element_text(size = 12),  # text size
    legend.position = "none"  # hide legend
  )

president_age_box_plot

```

<br>

# What did I learn?

In this week's *tidy_tuesday*, I learned **a new theme for markdown files** and **a new color pallete**. Both required me to install packages and load libraries, but it was worth it of course. The plot colors are very nice on the eyes, I may use it for my own data sometime soon. The new markdown theme presents the code in a very 'non-busy' format. I did want to learn how to make a new plot but was actually really curious as to what these results were going to look like! I will admit that I failed to animate this plot and eventually gave up but will do my research and come back with a good animated plot next week :)
