---
title: "class_9a"
author: "Brandon Brenes"
date: "2024-10-22"
output: html_document
---

##Functions! 
Today we're working on functions. Why use functions?
-Anytime you have a task that you want to use more than once
-Keeps your script clean
-Avoids mistakes with copy and paste
-Minimizes places where you need to update your code
```{r,message=FALSE,warning=FALSE}

#libraries

library(tidyverse)
library(palmerpenguins)
library(PNWColors)
library(ggplot2)
```


```{r,message=FALSE,warning=FALSE}

#First set-up your script and create a dataframe of random numbers

df <- tibble(
  a = rnorm(10), # draws 10 random values from a normal distribution
  b = rnorm(10),
  c = rnorm(10),
  d = rnorm(10)
)
head(df)

#Anytime you have to do a task more then once... Let's use an example where we want to rescale data in multiple columns (value - min/(max - min))...

#mutating preexisting column to make rescaled and  copy and paste the same code for each column without making any mistakes

df<-df %>% 
  mutate(a = (a-min(a, na.rm = TRUE))/(max(a, na.rm = TRUE)-min(a, na.rm = TRUE)),
         b = (b-min(b, na.rm = TRUE))/(max(b, na.rm = TRUE)-min(b, na.rm = TRUE)),
         c = (c-min(c, na.rm = TRUE))/(max(c, na.rm = TRUE)-min(c, na.rm = TRUE)),
         d = (d-min(d, na.rm = TRUE))/(max(d, na.rm = TRUE)-min(d, na.rm = TRUE)))

head(df)

#but copy and pasting can create problems and there may be typos so....

#we can write a function for this!
```

```{r,message=FALSE,warning=FALSE}

#ta da...our rescale function using the x as the input and using function() to create a function names rescale01

rescale01 <- function(x) {
  value<-(x-min(x, na.rm = TRUE))/(max(x, na.rm = TRUE)-min(x, na.rm = TRUE))
  return(value)
}

#now use the clean function - rescale01 - and instead of x we're using our real input values

df %>%
  mutate(a = rescale01(a),
         b = rescale01(b),
         c = rescale01(c),
         d = rescale01(d))

#letters are the column names

```
###There are three key steps to creating a new function:

    -You need to pick a name for the function. Here I’ve used rescale01 because this function rescales a vector to lie between 0 and 1.

    -You list the inputs, or arguments, to the function inside function. Here we have just one argument. If we had more the call would look like function(x, y, z).

    -You place the code you have developed in body of the function, a { block that immediately follows function(...).

    return tells us what values we want returned when you run the function
```{r,message=FALSE,warning=FALSE}
rescale01 <- function(x) {
  value<-(x-min(x, na.rm = TRUE))/(max(x, na.rm = TRUE)-min(x, na.rm = TRUE))
  return(value)
}
```

Note: It is way easier to create a function when you already wrote the code rather than creating the function from scratch

```{r,message=FALSE,warning=FALSE}

farenheit_to_celsius<- function(temp_F) {
  temp_C <- (temp_F - 32) * 5 / 9
return(temp_C)}


celsius_to_kelvin <- function(temp_C) {
  temp_K<-temp_C + 273.15
return(temp_K)}

#yay we made our first function!
```


```{r,message=FALSE,warning=FALSE}
#Making plots into a function

#Let's say you have a specifically formatted plot that you like and that you plan to use over and over again. By turning it into a function, you only have to code it once.

#Let's use our favorite penguin dataset and make a nice plot line plot.

pal<-pnw_palette("Lake",3, type = "discrete") 
#my color palette, choose three colors from 'lake', type= discrete means no gradients, relatively different colors.

#the bones of the plot...

ggplot(penguins, aes(x = body_mass_g, y = bill_length_mm, color = island))+#groups colored by island
  geom_point()+
  geom_smooth(method = "lm")+ # add a linear model
  scale_color_manual("Island", values=pal)+  # use pretty colors and another example of how to manually change the legend title for colors
  theme_bw()

```

We know that this works but we can expedite the process. Lets use our function...
```{r, message=FALSE,warning=FALSE}

#specify data is using x and y as the argument
#the function for ggplot needs something to recognize that its a column within that dataset
#use defaults, have the penguins also be the data frame its pulling from

myplot<-function(data = penguins, x, y) {
pal<-pnw_palette("Lake",3, type = "discrete") # my color palette 

ggplot(data, aes(x = {{x}}, y = {{y}} , color = island))+
  geom_point()+
  geom_smooth(method = "lm")+ # add a linear model
  scale_color_manual("Island", values=pal)+   # use pretty colors and change the legend title
  theme_bw()
}

myplot(x = body_mass_g, y = flipper_length_mm)

```

Imagine you want a variable to be equal to a certain value if a condition is met. This is a typical problem that requires the if ... else ... construct. For instance:
```{r,message=FALSE,warning=FALSE}
a <- 4
b <- 5
#Suppose that if a > b then f should be = to 20, else f should be equal to 10. Using if/else we:
if (a > b) { # my question
  f <- 20 # if it is true give me answer 1
    } else { # else give me answer 2
  f <- 10
}

f

```
Now lets apply this 'if-else' to our plot
```{r,message=FALSE,warning=FALSE}


myplot<-function(data = penguins, x, y, lines=TRUE ){ # add new argument for lines
pal<-pnw_palette("Lake",3, type = "discrete") # my color palette 

if(lines==TRUE){
ggplot(data, aes(x = {{x}}, y = {{y}} , color = island))+
  geom_point()+
  geom_smooth(method = "lm")+ # add a linear model
  scale_color_manual("Island", values=pal)+   # use pretty colors and change the legend title
  theme_bw()
}
else{
ggplot(data, aes(x = {{x}}, y = {{y}} , color = island))+
  geom_point()+
  scale_color_manual("Island", values=pal)+   # use pretty colors and change the legend title
  theme_bw()
}
}

myplot(x = body_mass_g, y = flipper_length_mm)

```

