---
title: "Class 8b"
author: "Brandon Brenes"
date: "2024-10-22"
output:
  html_document:
    toc: TRUE
    toc_float: TRUE
    theme: darkly
---

<br>

## Outline of class

#### Advanced plotting:
- {patchwork} - bringing plots together
- {ggrepel} - repel your labels on your plots
- {gganimate} - add some animations to yourplots
- {magick} - photo processing
    
<br>
    
#### Load libraries
```{r, message=FALSE,warning=FALSE}
#libraries
library(tidyverse)
library(here)
library(patchwork)
library(ggrepel)
library(gganimate)
library(magick)
library(palmerpenguins)
```

<br>

### patchwork
Easily brings the plots together with simple code, cal plot and gggrid also does this but patchwork is the easiest

<br>

#### How to use patchwork...

##### Plots being used:
```{r,message=FALSE,warning=FALSE}

# plot 1
p1<-penguins %>%
  ggplot(aes(x = body_mass_g, 
             y = bill_length_mm, 
             color = species))+
  geom_point()
p1

```

```{r,message=FALSE,warning=FALSE}

# plot 2
p2<-penguins %>%
  ggplot(aes(x = sex, 
             y = body_mass_g, 
             color = species))+
  geom_jitter(width = 0.2)
p2

```

<br>

#### Sticking plots together...

```{r,message=FALSE,warning=FALSE}
#use patchwork to add the plots together

p1+p2
#wow that actually worked!

```
<br>

#### Stack them ontop and label them...
```{r,message=FALSE,warning=FALSE}
p1/p2 +#stack on top
  plot_layout(guides = 'collect')+ #legend labels
  plot_annotation(tag_levels = 'A') #capital letters tagged

```


<br>

### ggrepel

We are using mtcars preloaded data set

```{r,message=FALSE,warning=FALSE}
 

ggplot(mtcars, aes(x = wt, 
                   y = mpg, 
                   label = rownames(mtcars)))+ #label them as whatever the row name is ex) cadillac
  geom_text_repel() + # repel the labels to see all of them, makes easier to read
  geom_point(color = 'red')

#instead of 'geom_text_repel' you could use 'geom_label_repel' for a different look

```


<br>

### gganimate
Useful when not distracting. Create animations like gifs.

##### you can show a transtiion between two states:
```{r,message=FALSE,warning=FALSE}
penguins %>%
ggplot(aes(x = body_mass_g, 
            y = bill_depth_mm, 
            color = species)) +
  geom_point() +
  transition_states(
     year, # what are we animating by
     transition_length = 2, #The relative length of the transition.
     state_length = 1 # The length of the pause between transitions
   )+
  ease_aes("sine-in-out") + #so many different animation styles ex) bounce-in-out
  labs(title = 'Year: {closest_state}')
  
  #add label to see what year it is, it reports the closest state, in this case state=year. must use brackets
  
# save animation #
#anim_save(here("Week_08","output","mypengiungif.gif"))
```


<br>

### magick
Advanced image processing. Helpful if you need to convert photos to gray scale, binary for an analysis, manipulate them in anyway, or putting images on top of your plots!

<br>

#### example:
```{r,message=FALSE,warning=FALSE}
penguin<-image_read("https://pngimg.com/uploads/penguin/pinguin_PNG9.png")#image_read reads the image in
#you can do if it it's on your desktop or output folder, use file path used in ggsave
penguin
```

<br>

```{r, include=FALSE}
penguins %>% #use this plot as an example
ggplot(aes(x = body_mass_g, 
            y = bill_depth_mm, 
            color = species)) +
  geom_point()  #ggsave(here("Week_08","output","penguinplot.png"))
```

<br>

#### Penguin cartoon on top of plot

```{r,message=FALSE,warning=FALSE}
#whatever you want to use as the background you would 'read in' first.

#penguin plot will be first since we used image_read...
penplot<-image_read(here("Week_08","output","penguinplot.png"))
#cartoon penguin ontop, using image_compos()
out <- image_composite(penplot, penguin, offset = "+70+30")

out #view

```

<br>

#### Penguin gif on top of plot

```{r,message=FALSE,warning=FALSE}

# read in a penguin gif first #

pengif<-image_read("https://media3.giphy.com/media/H4uE6w9G1uK4M/giphy.gif")


# how to display gif on plot #

outgif <- image_composite(penplot, pengif, gravity = "center")

#penplot first then pengif
#gravity = center puts it right in middle of plot
#image_composite() on penplot


# extra step using image_animate #

animation <- image_animate(outgif, fps = 10, optimize = TRUE)
#low fps to save space, optimized it to smoothin it
#customize the gif settings

## save animation by using anim_save(here("Week_08","output","mypengiunplotgif.gif"))

# view #
animation

```

