---
title: "class 8a"
author: "Brandon Brenes"
date: "2024-10-15"
output:
  html_document:
    toc: TRUE
    toc_float: TRUE
    theme: darkly
---

### Descritpion:
We are learning {strings}...


    1.Using {stringr} to clean up strings (part of tidyverse)
    2.Intro to regex (regular expressions
    3. Using {tidytext} for text mining/analysis
    4. Make a wordcloud

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(here)
library(tidyverse)
library(tidytext)
library(wordcloud2)
library(janeaustenr)
```

### Intro

There are 4 basic families of functions in the {stringr} package:

Manipulation: these functions allow you to manipulate individual characters within the strings in character vectors.

Whitespace tools: to add, remove, and manipulate whitespace.

Locale sensitive operations: whose operations will vary from locale to locale.

Pattern matching functions: These recognize four engines of pattern description. The most common is regular expressions, but there are three other tools.


### examples of strings...
```{r}

words<-"This is a string"
words

words_vector<-c("Apples", "Bananas","Oranges")
words_vector

```

### examples of manipulation...

```{r}
paste("High temp", "Low pH")

paste("High temp", "Low pH", sep = "-") #Add a dash in between the words

paste0("High temp", "Low pH") #Remove the space in between the words

shapes <- c("Square", "Circle", "Triangle")
paste("My favorite shape is a", shapes) #Working with vectors 

two_cities <- c("best", "worst")
paste("It was the", two_cities, "of times.")

#Lets say you wanna know how long a string is...

shapes # vector of shapes
str_length(shapes) # how many letters are in each word?

#Let's say you want to extract specific characters. Do you work with sequence data? This could be super useful to exact specific bases in a sequence.

seq_data<-c("ATCCCGTC")
str_sub(seq_data, start = 2, end = 4) # extract the 2nd to 4th AA

#You can also modify strings

str_sub(seq_data, start = 3, end = 3) <- "A" # add an A in the 3rd position
seq_data

#You can also duplicate patterns in your strings. Here I am duplicating it 2 and 3 times

str_dup(seq_data, times = c(2, 3)) # times is the number of times to duplicate each string

```

### Whitespace

```{r}

#Say you have a column and you did not copy and paste your treatments like you learned in the first week of class. You now have some words with extra white spaces and R thinks its an entirely new word. Here is how to deal with that...

badtreatments<-c("High", " High", "High ", "Low", "Low")
badtreatments

#Remove white space

str_trim(badtreatments) # this removes both

#You can also just remove from one side or the other

str_trim(badtreatments, side = "left") # this removes left

#The opposite of str_trim is str_pad, to add white space to either side

str_pad(badtreatments, 5, side = "right") # add a white space to the right side af

str_pad(badtreatments, 5, side = "right", pad = "1") # add a 1 to the right side after the 5th character
```
### locale sensitive
```{r}

#Make everything upper case

x<-"I love R!"
str_to_upper(x)

#Make it title case (Cap first letter of each word)

str_to_title(x)

```

### pattern matching
{stringr} has functions to view, detect, locate, extract, match, replace, and split strings based on specific patterns
```{r}
data<-c("AAA", "TATA", "CTAG", "GCTT")
#visualize
str_view(data, pattern = "A")
#detect
str_detect(data, pattern = "AT")

#Locate a pattern

str_locate(data, pattern = "AT")
```
### regex: regular expressions
There are several types of regular expressions:

  **Metacharacters
   ** Sequences
   ** Quantifiers
    **Character classes
    **POSIX character classes (Portable Operating System Interface)
    
**Metacharacters:
Metacharacters: The simplest form of regular expressions are those that match a single character. Most characters, including all letters and digits, are regular expressions that match themselves. For a language like R, there are some special characters that have reserved meaning and they are referred to as ‘Metacharacters”. The metacharacters in Extended Regular Expressions (EREs) are:

. \ | ( ) [ { $ * + ?

```{r}
#Let's say that you have the following set of strings...

vals<-c("a.b", "b.c","c.d")

#And you want to replace all the "." with a space. Here is how you would do it:

#string, pattern, replace
str_replace(vals, "\\.", " ")


#Let's say we had multiple "." in our character vector

vals<-c("a.b.c", "b.c.d","c.d.e")
#string, pattern, replace
str_replace(vals, "\\.", " ")

## [1] "a b.c" "b c.d" "c d.e"

#str_replace only replaces the first instance. Let's try str_replace_all()

#string, pattern, replace
str_replace_all(vals, "\\.", " ") #replaces alllllll of them

```
Sequences, as the name suggests refers to the sequences of characters which can match. We have shorthand versions (or anchors) for commonly used sequences in R:

see slides for examps

#Let's subset the vector to only keep strings with digits
```{r}
val2<-c("test 123", "test 456", "test")
str_subset(val2, "\\d")
```
### character class
A character class or character set is a list of characters enclosed by square brackets [ ]. Character sets are used to match only one of the different characters. For example, the regex character class [aA] matches any lower case letter a or any upper case letter A

see slides for examps

```{r}
#count the number of lowercase vowels in each string
str_count(val2, "[aeiou]") #you'll get "1"s because only one vowel in the word 'test'
```

### quantifiiers
Example: find the phone numbers
```{r}
strings<-c("550-153-7578",
         "banana",
         "435.114.7586",
         "home: 672-442-6739")



#Make a regex that finds all the strings that contain a phone number. We know there is a specific pattern (3 numbers, 3 numbers, 4 numbers and it can have either a "." or "-" to separate them). Let's also say we know that the first number cannot be a 1

phone <- "([2-9][0-9]{2})[- .]([0-9]{3})[- .]([0-9]{4})" #the{} is the number of digits that have that range

# Which strings contain phone numbers?
test<-str_subset(strings, phone)

test %>% 
  str_replace_all(pattern = "\\.", replacement = "-") %>% 
  str_replace_all(pattern = "[a-zA-Z]|\\:", replacement = "") %>% 
  str_trim()

```

```{r}
original_books <- austen_books() %>% # get all of Jane Austen's books
  group_by(book) %>%
  mutate(line = row_number(), # find every line
         chapter = cumsum(str_detect(text, regex("^chapter [\\divxlc]", # count the chapters (starts with the word chapter followed by a digit or roman numeral), the "^" means 'start with'
                                                 ignore_case = TRUE)))) %>% #ignore lower or uppercase
  ungroup() # ungroup it so we have a dataframe again
# don't try to view the entire thing... its >73000 lines...
head(original_books)

#Because we are interest in text mining, we will want to clean this so that there is only one word per row so its tidy. In tidytext each word is refered to as a token. The function to unnest the data so that its only one word per row is unnest_tokens().

tidy_books <- original_books %>%
  unnest_tokens(output = word, input = text) # add a column named word, with the input as the text column
head(tidy_books) # there are now >725,000 rows. Don't view the entire thing!


#lot of notes on tidytext
```

